---
title: "Customer Personality Analysis"
author: "Aashay Sharma"
date: "16/11/2021"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Introduction 
Customer Personality analysis is a detailed analysis done by a brand or company to find an ideal set of customers. This helps the company to make changes in their product and marketing in a more efficient way. Promoting a product to a customer who is more likely to buy the product is better than promoting it to a random person. 

Customer Personality analysis is done by clustering as it is a un-supervised problem. 

```{r,warning = FALSE, results= FALSE}
library(dplyr)
library(ggplot2)
library(cluster)
library(gower)
library(Rtsne)
library(stringr)
library(corrgram)
```

# Data 
```{r, warning = FALSE}
data <- read.csv2("../input/customer-personality-analysis/marketing_campaign.csv", sep = "\t")

sum(is.na(data))

data <- na.omit(data)
```
As there are very few NA items we will just omit them.

```{r, warning=FALSE}
str(data)
```
The data is mixed type data (ie both Categorical and Continous values). Euclidean distance is commonly used for clustering algorithms like KMeans and hierarchical clustering but it is useful when we only have to deal with continous values. Thus here we will use gower distance as the distance metric.

Gower distance is computed as the average of partial dissimilarities across individuals. Each partial dissimilarity (and thus Gower distance) ranges in [0 1].

# Data Cleaning 
To work with the data we need to alter some variables and add a few new variables based on the existing ones. 

1. Year_Birth 
Changing it to the Age variable 
```{r, warning=FALSE}
data$Age <- 2014 - data$Year_Birth
```

2. Education 
It has too many categories this can be converted to UG and PG categories.
```{r,warning=FALSE}
data$Education[data$Education == "2n Cycle"] = "UG"
data$Education[data$Education == "Basic"] = "UG"
data$Education[data$Education == "Graduation"] = "PG"
data$Education[data$Education == "Master"] = "PG"
data$Education[data$Education == "PhD"] = "PG"
```

3. Marital_Status 
It also has many categories which can be converted to Single and Couple categories.

```{r, warning=FALSE}
data$Marital_Status[data$Marital_Status == "Divorced"] = "Single"
data$Marital_Status[data$Marital_Status == "Absurd"] = "Single"
data$Marital_Status[data$Marital_Status == "YOLO"] = "Single"
data$Marital_Status[data$Marital_Status == "Widow"] = "Single"
data$Marital_Status[data$Marital_Status == "Together"] = "Couple"
data$Marital_Status[data$Marital_Status == "Married"] = "Couple"
data$Marital_Status[data$Marital_Status == "Alone"] = "Single"
```

4. Dt_Customer
The customer association date with the company can be changed to a seniority level.
```{r,warning=FALSE}
data$Customer_year <- str_sub(data$Dt_Customer,-4)
data$Customer_year <- as.numeric(data$Customer_year)
data$Customer_Seniority <- 2014 - data$Customer_year
```

5. Kid home and Teen Home 
I feel that the keed home and teen home variables can collectively define the child variable.

```{r,warning=FALSE}
data$Child <- data$Kidhome + data$Teenhome
```

6. Mnt products 
The amount spend of products can be added together to get the total amount spent which will be a more useful variable.

```{r, warning=FALSE}
data$Amt_Spent <- data$MntWines + data$MntFishProducts + data$MntFruits + data$MntGoldProds + data$MntMeatProducts + data$MntSweetProducts
```

7. Purchases made 
We can get the total number of purchases by adding the different purchases variables.
```{r, warning=FALSE}
data$Num_Purchases_made <- data$NumWebPurchases + data$NumCatalogPurchases + data$NumStorePurchases
```

Thus we are left with 12 variables that can be used to create the distance matrix.
```{r,warning=FALSE}
data <- data[c(1,30,3,4,5,33,32,9,34,35,16,20)]
```

# EDA 

### Corrgram : 

```{r,warning=FALSE}
data2 <- data
data2$Education <- unclass(as.factor(data2$Education))
data2$Marital_Status <- unclass(as.factor(data2$Marital_Status))
data2$Education <- as.numeric(data2$Education)
data2$Marital_Status <- as.numeric(data2$Marital_Status)
corrgram(data2[-c(1)], order=TRUE, lower.panel=panel.shade, upper.panel=NULL, text.panel=panel.txt, main="Customer Data")

```
Strong Correlation between Income, Amount Spent and Purchase, As well as there is slight trend between Education and Income(PG:1, UG:2) so this tells us that people with higher studies earn more. 
1. Age

```{r,warning=FALSE}
age_plot <- ggplot(data = data, aes(Age))
age_plot + geom_density()
```

2. Education 
```{r,warning=FALSE}
edu_plot1 <- ggplot(data = data, aes(Education))
edu_plot1 + geom_histogram(stat = "count")
```

```{r,warning=FALSE}
edu_plot2<- ggplot(data = data, aes(Age, fill = Education))
edu_plot2 + geom_histogram() 
```

3. Marital_Status 
```{r,warning=FALSE}
marital_plot <- ggplot(data = data, aes(Marital_Status, fill = Marital_Status))
marital_plot + geom_histogram(stat = "count")
```

4. Income & Amount Spent 
```{r,warning=FALSE}
require(scales)
inc_plt <- ggplot(data = data, aes(Income, Amt_Spent))
inc_plt + geom_point(alpha = 0.5, color = "blue") + scale_x_continuous(labels = comma)
```

5. Child 
```{r,warning=FALSE}
a <- ggplot(data = data,  aes(Child, fill = Marital_Status))
a + geom_histogram(position = "dodge")
```

6. Customer Seniority & Amount Spent 
```{r, warning = FALSE}
cust_plt <- ggplot(data = data, aes(as.factor(Customer_Seniority), Amt_Spent, fill = Customer_Seniority))

cust_plt + geom_boxplot(color = "black")
```

# Customer Analysis : 

## Analysing customers based on their income, seniority, age and amount spent.

```{r, warning = FALSE}
gower_dist2 <- daisy(data[c("Income", "Amt_Spent", "Customer_Seniority", "Age")],
                    metric = "gower")

summary(gower_dist2)

gower_mat2 <- as.matrix(gower_dist2)
```

MOST SIMILAR PAIR 
```{r,warning=FALSE}
data[
  which(gower_mat2 == min(gower_mat2[gower_mat2 != min(gower_mat2)]),
        arr.ind = TRUE)[1, ], ]
```

MOST DISSIMILAR PAIR
```{r, warning=FALSE}
data[
  which(gower_mat2 == max(gower_mat2[gower_mat2 != max(gower_mat2)]),
        arr.ind = TRUE)[1, ], ]
```

FINDING NUMBER OF CLUSTERS 
```{r, warning=FALSE}
sil_width2 <- c(NA)

for(i in 2:10){
  
  pam_fit2 <- pam(gower_dist2,
                 diss = TRUE,
                 k = i)
  
  sil_width2[i] <- pam_fit2$silinfo$avg.width
  
}
```

```{r, warning=FALSE}
plot(1:10, sil_width2,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width2)
```
By silhouette method we can choose 4 clusters.

CLUSTER SUMMARY
```{r, warning=FALSE}
pam_fit2 <- pam(gower_dist2, diss = TRUE, k = 4)

pam_results2 <- data %>%
  dplyr::select(-c(ID, Child, NumWebVisitsMonth, Num_Purchases_made, NumDealsPurchases, Education, Marital_Status)) %>%
  mutate(cluster = pam_fit2$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results2$the_summary
```
We can interpret from the cluster summary that 4 types of customers here are : 

### 1. Cluster 1 : Avg Income | Senior Customers | Spending Avg Amount 
### 2. Cluster 2 : Avg Income | New Customers | Avg spending less than cluster 1 customers
### 3. Cluster 3 : High Income | Fairly Senior Customers | High Avg Spending 
### 4. Cluster 4 : Below Avg Income | Fairly Senior Customers | Low Average Spending 

VISUALISING CLUSTERS
```{r, warning=FALSE}
tsne_obj2 <- Rtsne(gower_dist2, is_distance = TRUE)

tsne_data2 <- tsne_obj2$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit2$clustering),
         name = data$ID)

ggplot(aes(x = X, y = Y), data = tsne_data2) +
  geom_point(aes(color = cluster))
```


## Analysing customers based on Item Purchased, Amount Spent, Number of Child and Marital Status.

```{r,warning=FALSE}
data2 <- data

data2$Education <- unclass(as.factor(data2$Education))

data2$Marital_Status <- unclass(as.factor(data2$Marital_Status))

data2$Education <- as.numeric(data2$Education)

data2$Marital_Status <- as.numeric(data2$Marital_Status)

data_new <- read.csv2("../input/customer-personality-analysis/marketing_campaign.csv", sep = "\t")
data_new <- na.omit(data_new)

data3 <- data2
data3$AmtWines <- data_new$MntWines
data3$AmtFruits <- data_new$MntFruits
data3$AmtFish <- data_new$MntFishProducts
data3$AmtMeat <- data_new$MntMeatProducts
data3$AmtSweet <- data_new$MntSweetProducts
data3$AmtGold <- data_new$MntGoldProds
data3$Education <- as.factor(data3$Education)
data3$Marital_Status <- as.factor(data3$Marital_Status)
```

```{r, warning = FALSE}
gower_dist3 <- daisy(data3[c("Income", "Child", "Num_Purchases_made", "Marital_Status", "AmtWines", "AmtFruits", "AmtFish", "AmtMeat", "AmtSweet", "AmtGold", "NumDealsPurchases", "Amt_Spent", "Recency")],
                    metric = "gower")

summary(gower_dist3)

gower_mat3 <- as.matrix(gower_dist3)
```

MOST SIMILAR PAIR 
```{r,warning=FALSE}
data3[
  which(gower_mat3 == min(gower_mat3[gower_mat3 != min(gower_mat3)]),
        arr.ind = TRUE)[1, ], ]
```

MOST DISSIMILAR PAIR
```{r, warning=FALSE}
data3[
  which(gower_mat3 == max(gower_mat3[gower_mat3 != max(gower_mat3)]),
        arr.ind = TRUE)[1, ], ]
```

FINDING NUMBER OF CLUSTERS 
```{r, warning=FALSE}
sil_width3 <- c(NA)

for(i in 2:10){
  
  pam_fit3 <- pam(gower_dist3,
                 diss = TRUE,
                 k = i)
  
  sil_width3[i] <- pam_fit3$silinfo$avg.width
  
}
```

```{r, warning=FALSE}
plot(1:10, sil_width3,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width3)
```
By silhouette method we can choose 4 clusters.

CLUSTER SUMMARY
```{r, warning=FALSE}
pam_fit3 <- pam(gower_dist3, diss = TRUE, k = 3)

pam_results3 <- data3 %>%
  dplyr::select(-c(ID, NumWebVisitsMonth, Age, Customer_Seniority)) %>%
  mutate(cluster = pam_fit3$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results3$the_summary
```
We can interpret from the cluster summary that 3 types of customers here are : 

### 1. Cluster 1 : Couple or Not Single people | More post graduates | Higher than average income | Highest Buyer of Wine & Meat | Gold purchase also above avg | Overall Purchase and Amount Spent also Highest 

### 2. Cluster 2 : Singles | Post Graduates | Average Income | Below Avg Amount Spent | Greater Deal Purchases than cluster one customers 

### 3. Cluster 3 : Couples | Mostly Post Graduates (But also includes relatively higher number of Undergrads compared to other clusters) | Below Avg Income | Low Amount Spent | More Deal focused Purchases | Would prefer essentials over lesuire (Gold and Sweets)

VISUALISING CLUSTERS
```{r, warning=FALSE}
tsne_obj3 <- Rtsne(gower_dist3, is_distance = TRUE)

tsne_data3 <- tsne_obj3$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit3$clustering),
         name = data3$ID)

ggplot(aes(x = X, y = Y), data = tsne_data3) +
  geom_point(aes(color = cluster))
```
# General Conclusions : 

## 1. New & Old Customers with Avg Income and fairly avg amount spent should be focused on more, Better advertising and deals should be provided to them.

## 2. If a new Non discounted expensive item will be up for sale ads should be targeted better to customers with high spending nature & income.

## 3. Customers with low spending natures and low incomes should be targeted with flash deals and discounts on essentials like fish, fruits & meat to keep them connected with the company.

## 4. Customers tend to spend & purchase more on wine and meat so this products can be worked upon (Like more deals, better more variable products etc) to benefit the company.

## 5. Senior Customers should be connected with company in some or other ways like memberships or specific deals to them this will increase their relationship with company.

