---
title: "Customer Personality Analysis"
author: "Group 7 - Isaac Yang, Pengxiang Chen, Raihan Putra, Zhiqi Rui"
date: "2021-12-09"
output:
  html_document: default
  pdf_document: default
---

```{r, echo = FALSE, warning = FALSE}
rm(list = ls())

library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

# Introduction 
Customer personality analysis is a detailed analysis done by a brand or company to find an ideal set of customers. This helps the company to make changes in their product and marketing in a more efficient way. Promoting a product to a customer who is more likely to buy the product is better than promoting it to a random person.

Customer personality analysis is done by clustering as it is an unsupervised problem. 

```{r, warning = FALSE, results = FALSE}
library(dplyr)
library(ggplot2)
library(cluster)
library(gower)
library(Rtsne)
library(stringr)
library(corrgram)
```

# Data 
```{r, warning = FALSE}
data <- read.csv2("../data/marketing_campaign.csv", sep = "\t")

sum(is.na(data))
data <- na.omit(data)
```
As there are very few NA items we will just omit them.

```{r, warning = FALSE}
str(data)
```
The data is mixed type data (i.e. both categorical and continuous values). Euclidean distance is commonly used for clustering algorithms like KMeans and hierarchical clustering but it is useful when we only have to deal with continuous values. Thus, we will use Gower distance as the distance metric.

Gower distance is computed as the average of partial dissimilarities across individuals. Each partial dissimilarity (and thus Gower distance) ranges in [0 1].

# Data Cleaning 
To work with the data we need to alter some variables and add a few new variables based on the existing ones. 

1. Year_Birth, changing it to the Age variable.
```{r, warning = FALSE}
data$Age <- 2014 - data$Year_Birth
```

2. Education, it has too many categories this can be converted to UG and PG categories.
```{r, warning = FALSE}
data$Education[data$Education == "2n Cycle"] = "UG"
data$Education[data$Education == "Basic"] = "UG"
data$Education[data$Education == "Graduation"] = "PG"
data$Education[data$Education == "Master"] = "PG"
data$Education[data$Education == "PhD"] = "PG"
```

3. Marital_Status, it also has many categories which can be converted to Single and Couple categories.
```{r, warning = FALSE}
data$Marital_Status[data$Marital_Status == "Divorced"] = "Single"
data$Marital_Status[data$Marital_Status == "Absurd"] = "Single"
data$Marital_Status[data$Marital_Status == "YOLO"] = "Single"
data$Marital_Status[data$Marital_Status == "Widow"] = "Single"
data$Marital_Status[data$Marital_Status == "Together"] = "Couple"
data$Marital_Status[data$Marital_Status == "Married"] = "Couple"
data$Marital_Status[data$Marital_Status == "Alone"] = "Single"
```

4. Dt_Customer, the customer association date with the company can be changed to a seniority level.
```{r, warning = FALSE}
data$Customer_year <- str_sub(data$Dt_Customer, -4)
data$Customer_year <- as.numeric(data$Customer_year)
data$Customer_Seniority <- 2014 - data$Customer_year
```

5. Kidhome and Teenhome, we feel that the Kidhome and Teenhome variables can collectively define the child variable.
```{r, warning = FALSE}
data$Child <- data$Kidhome + data$Teenhome
```

6. Mnt products, the amount spend of products can be added together to get the total amount spent which will be a more useful variable.
```{r, warning = FALSE}
data$Amt_Spent <-
  data$MntWines + data$MntFishProducts + data$MntFruits + data$MntGoldProds + data$MntMeatProducts + data$MntSweetProducts
```

7. Num_Purchases_made, we can get the total number of purchases by adding the different purchases variables.
```{r, warning = FALSE}
data$Num_Purchases_made <-
  data$NumWebPurchases + data$NumCatalogPurchases + data$NumStorePurchases
```

8. Thus, we are left with 12 variables that can be used to create the distance matrix.
```{r, warning = FALSE}
data <- data[c(1, 30, 3, 4, 5, 33, 32, 9, 34, 35, 16, 20)]
data <- data[data$Income<4e05,]
```

# Exploratory Data Analysis
## Corrgram 

```{r, warning = FALSE}
data2 <- data
data2$Education <- unclass(as.factor(data2$Education))
data2$Marital_Status <- unclass(as.factor(data2$Marital_Status))
data2$Education <- as.numeric(data2$Education)
data2$Marital_Status <- as.numeric(data2$Marital_Status)
corrgram(
  data2[-c(1)],
  order = TRUE,
  lower.panel = panel.shade,
  upper.panel = NULL,
  text.panel = panel.txt,
  main = "Customer Data"
)
```

Strong Correlation between Income, Amount Spent and Purchase, as well as slight trend between Education and Income (PG: 1, UG: 2) tells us that people with higher studies earn more.

1. Age
```{r, warning = FALSE}
age_plot <- ggplot(data = data, aes(Age))
age_plot + geom_density()
```

2. Education 
```{r, warning = FALSE}
edu_plot1 <- ggplot(data = data, aes(Education))
edu_plot1 + geom_histogram(stat = "count")
```

```{r, warning = FALSE}
edu_plot2 <- ggplot(data = data, aes(Age, fill = Education))
edu_plot2 + geom_histogram() 
```

3. Marital Status 
```{r, warning = FALSE}
marital_plot <-
  ggplot(data = data, aes(Marital_Status, fill = Marital_Status))
marital_plot + geom_histogram(stat = "count")
```

4. Income & Amount Spent 
```{r, warning = FALSE}
require(scales)
inc_plt <- ggplot(data = data, aes(Income, Amt_Spent))
inc_plt + geom_point(alpha = 0.5, color = "blue") + scale_x_continuous(labels = comma)
```

5. Child 
```{r, warning = FALSE}
a <- ggplot(data = data,  aes(Child, fill = Marital_Status))
a + geom_histogram(position = "dodge")
```

6. Customer Seniority & Amount Spent 
```{r, warning = FALSE}
cust_plt <-
  ggplot(data = data, aes(as.factor(Customer_Seniority), Amt_Spent, fill = Customer_Seniority))

cust_plt + geom_boxplot(color = "black")
```

# Analysis
## Analysis based on income, seniority, age, and amount spent.

```{r, warning = FALSE}
gower_dist2 <-
  daisy(data[c("Income", "Amt_Spent", "Customer_Seniority", "Age")],
        metric = "gower")

summary(gower_dist2)

gower_mat2 <- as.matrix(gower_dist2)
```

### Most similar pair
```{r, warning = FALSE}
data[which(gower_mat2 == min(gower_mat2[gower_mat2 != min(gower_mat2)]),
           arr.ind = TRUE)[1,],]
```

### Most dissimilar pair
```{r, warning = FALSE}
data[which(gower_mat2 == max(gower_mat2[gower_mat2 != max(gower_mat2)]),
           arr.ind = TRUE)[1,],]
```

### Finding number of clusters
```{r, warning = FALSE}
sil_width2 <- c(NA)

for (i in 2:10) {
  pam_fit2 <- pam(gower_dist2,
                  diss = TRUE,
                  k = i)
  
  sil_width2[i] <- pam_fit2$silinfo$avg.width
  
}
```

```{r, warning = FALSE}
plot(1:10, sil_width2,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width2)
```

By silhouette method we can choose 4 clusters.

### Cluster summary
```{r, warning = FALSE}
pam_fit2 <- pam(gower_dist2, diss = TRUE, k = 4)

pam_results2 <- data %>%
  dplyr::select(
    -c(
      ID,
      Child,
      NumWebVisitsMonth,
      Num_Purchases_made,
      NumDealsPurchases,
      Education,
      Marital_Status
    )
  ) %>%
  mutate(cluster = pam_fit2$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results2$the_summary
```
We can interpret from the cluster summary that 4 types of customers here are: 

* Cluster 1: Avg Income | Senior Customers | Spending Avg Amount
* Cluster 2: Avg Income | New Customers | Avg Spending less than Cluster 1 customers
* Cluster 3: High Income | Fairly Senior Customers | High Avg Spending 
* Cluster 4: Below Avg Income | Fairly Senior Customers | Low Avg Spending 

### Visualizing cluster
```{r, warning = FALSE}
tsne_obj2 <- Rtsne(gower_dist2, is_distance = TRUE)

tsne_data2 <- tsne_obj2$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit2$clustering),
         name = data$ID)

ggplot(aes(x = X, y = Y), data = tsne_data2) +
  geom_point(aes(color = cluster))
```

## Analysis based on item purchased, amount spent, number of child, and marital status.
```{r, warning = FALSE}
data2 <- data

data2$Education <- unclass(as.factor(data2$Education))

data2$Marital_Status <- unclass(as.factor(data2$Marital_Status))

data2$Education <- as.numeric(data2$Education)

data2$Marital_Status <- as.numeric(data2$Marital_Status)

data_new <- read.csv2("../data/marketing_campaign.csv", sep = "\t")
data_new <- na.omit(data_new)
data_new <- data_new[data_new$Income<4e05,]

data3 <- data2
data3$AmtWines <- data_new$MntWines
data3$AmtFruits <- data_new$MntFruits
data3$AmtFish <- data_new$MntFishProducts
data3$AmtMeat <- data_new$MntMeatProducts
data3$AmtSweet <- data_new$MntSweetProducts
data3$AmtGold <- data_new$MntGoldProds
data3$Education <- as.factor(data3$Education)
data3$Marital_Status <- as.factor(data3$Marital_Status)
```

```{r, warning = FALSE}
gower_dist3 <-
  daisy(data3[c(
    "Income",
    "Child",
    "Num_Purchases_made",
    "Marital_Status",
    "AmtWines",
    "AmtFruits",
    "AmtFish",
    "AmtMeat",
    "AmtSweet",
    "AmtGold",
    "NumDealsPurchases",
    "Amt_Spent",
    "Recency"
  )],
  metric = "gower")

summary(gower_dist3)

gower_mat3 <- as.matrix(gower_dist3)
```

### Most similar pair
```{r, warning = FALSE}
data3[which(gower_mat3 == min(gower_mat3[gower_mat3 != min(gower_mat3)]),
            arr.ind = TRUE)[1, ], ]
```

### Most dissimilar pair
```{r, warning = FALSE}
data3[which(gower_mat3 == max(gower_mat3[gower_mat3 != max(gower_mat3)]),
            arr.ind = TRUE)[1, ], ]
```

### Finding number of clusters
```{r, warning = FALSE}
sil_width3 <- c(NA)

for (i in 2:10) {
  pam_fit3 <- pam(gower_dist3,
                  diss = TRUE,
                  k = i)
  
  sil_width3[i] <- pam_fit3$silinfo$avg.width
  
}
```

```{r, warning = FALSE}
plot(1:10, sil_width3,
     xlab = "Number of clusters",
     ylab = "Silhouette Width")
lines(1:10, sil_width3)
```

By silhouette method we can choose 4 clusters.

### Cluster summary
```{r, warning = FALSE}
pam_fit3 <- pam(gower_dist3, diss = TRUE, k = 3)

pam_results3 <- data3 %>%
  dplyr::select(-c(ID, NumWebVisitsMonth, Age, Customer_Seniority)) %>%
  mutate(cluster = pam_fit3$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))

pam_results3$the_summary
```
We can interpret from the cluster summary that 3 types of customers here are: 

* Cluster 1: Couple or Not Single people | More Post Graduates | Higher than Avg Income | Highest Buyer of Wine & Meat | Gold Purchase also Above Avg | Overall Purchase and Amount Spent also Highest 

* Cluster 2: Singles | Post Graduates | Avg Income | Below Avg Amount Spent | Greater Deal Purchases than Cluster One Customers 

* Cluster 3: Couples | Mostly Post Graduates (But also includes relatively higher number of Undergrads compared to other clusters) | Below Avg Income | Low Amount Spent | More Deal focused Purchases | Would prefer essentials over leisure (Gold and Sweets)

### Visualizing cluster
```{r, warning = FALSE}
tsne_obj3 <- Rtsne(gower_dist3, is_distance = TRUE)

tsne_data3 <- tsne_obj3$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit3$clustering),
         name = data3$ID)

ggplot(aes(x = X, y = Y), data = tsne_data3) +
  geom_point(aes(color = cluster))
```

# Conclusion

1. New and old customers with avg income and fairly avg amount spent should be more focused on, provide better advertising and deals to them.
2. If a new non-discounted expensive item will be up for sale, ads should be targeted better to customers with high spending nature and income.
3. Customers with low spending natures and low incomes should be targeted with flash deals and discounts on essentials like fish, fruits, and meat to keep them engaged with the company.
4. Customers tend to spend and purchase more on wine and meat so this products can be worked upon such as more deals and more variability on the product to benefit the company.
5. Senior customers should be engaged with the company in other ways such as memberships or specific deals, this will increase improve relationship with company.
